---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "Chapter 17-20: Metaprogramming"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

# 18. Expressions

# Abstract Syntax Trees

## 1.

```
f(g(h))
1 + 2 + 3
(x + y) *z
```

## 2. 

```{r, eval = FALSE}
library(lobstr)

ast(f(g(h(i(1, 2, 3)))))
ast(f(1, g(2, h(3, i()))))
ast(f(g(1, 2), h(3, i(4, 5))))
```

## 3.

The first just uses backticks, which aren't quoted. In the second, `**` is an alias for `^`. In the third, right assign is an alias for left assign with the argument order swapped.

## 4.

`srcref` indicates an inlined expression within another, here the empty body of the created function.
We also see how the formals and body components are separated by the `function` function.

## 5.

The subsequent `elses` are nested within the previous `if`, since `if...else` is syntactic sugar for 
repeated calls of the `if` function.
```{r, eval = FALSE}
ast(
  if (x < 4) {
    y
  } else if (x < -2) {
    z
  } else if (q) {
    d
  } else {
    v
  }
)
```

# Parsing and Grammar

## 1.

It can contain both constants and calls. At the top level, it prints output explicitly.
```{r, eval = FALSE}
ast(f((1)))
ast(`(`(1 + 1))
```

## 2.


It should not be used for assignment.
```{r}
mean(x = 1:10)
x <- 1:10
```

## 3.

-4, because `^` has precedence over `-`.
```{r}
-2^2
```

## 4.

The outer `!` modifies the whole expression. 
The inner expression resolves to 1 because `FALSE` coerces to a numeric 0, and coercing 1 to logical and negating gives `FALSE`. So the result is `FALSE`.

```{r, eval = FALSE}
ast(!1 + !1)
```

## 5.

The chaining works because `<-` returns its 
right-hand side, and lazy evaluation ensures no name is evaluated before it is bound.

```{r}
x1 <- x2 <- x3 <- 0
```

## 6.

They beat unary but not binary operators.
```{r, eval = FALSE}
ast(x + y %+% z)
ast(x^y %+% z)
```

## 7. 

Sensibly, it throws an error.
```{r}
try(parse_expr("x + 1; y + 1"))
```

## 8. 

"Unexpected end of input."
```{r}
try(parse_expr("a +"))
```

## 9. 

It returns parsed lines of a given width. 
But the lifecycle is questioning!
```{r}
expr_text(lm(data = mtcars, formula = cyl ~ .))
```

## 10. 

To my surprise, it doesn't break.
```{r}

pairwise.t.test(Lorem <- ipsum <- dolor <- sit <- ametconsetetur <- sadipscing <- elitrsed <- diam <- nonumy <- eirmod <- tempor <- invidunt <- ut <- labore <- et <- dolore <- magna <- aliquyam <- eratsed <- diam <- voluptua. <- At <- vero <- eos <- et <- accusam <- et <- justo <- duo <- dolores <- et <- ea <- rebum. <- Stet <- clita <- kasd <- gubergrenno <- sea <- takimata <- sanctus <- est <- Lorem <- ipsum <- dolor <- sit <- amet <- aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa <- 1:100, g = 1:100 %/% 10)
```

# Expressions 

## 1.

Raw and complex, which have constructor functions and so can't be entered directly as constants.

## 2.

It turns into a call with the new first element as the outermost function, which often means the result is nonsensical.
```{r}
library(rlang)

x <- expr(read.csv("foo.csv", header = TRUE))
x[-1]
(x[-1])[-1]
```

## 3.

The first unquotes all call components. The second quotes `median`, deferring method dispatch. The third quotes `x`, so the generic delegates to `UseMethod` but `x` isn't substituted.
The fourth quotes both elements, leaving the call as entered.

```{r}
x <- 1:10

call2(median, x, na.rm = TRUE)
call2(expr(median), x, na.rm = TRUE)
call2(median, expr(x), na.rm = TRUE)
call2(expr(median), expr(x), na.rm = TRUE)
```

## 4.

`mean` passes `na.rm` via dots, and it is only matched in `mean.default`, not the `mean` generic, so the function can't standardize the call.

## 5.

`foo` is a function being called, not a tagged argument, so it needs no name, and indeed the `names` attribute only exists in the call object and will not impact the call when executed

```{r}
x <- expr(foo(x = 1))
names(x) <- c("x", "y")
x
```

## 6.

The arguments to `call2` appear in the order of the
code: test, if, else.
```{r}
greater <- call2(quote(`>`), quote(x), 1)
call2(quote(`if`), greater, "a", "b")
```

## AST Walking

I added the case for closures.
```{r}
expr_type <- function(x) {
  if (rlang::is_syntactic_literal(x)) {
    "constant"
  } else if (is.symbol(x)) {
    "symbol"
  } else if (is.call(x)) {
    "call"
  } else if (is.pairlist(x)) {
    "pairlist"
  } else if (is.integer(x) && "srcfile" %in% names(attributes(x))) {
    "closure"
  } else {
    typeof(x)
  }
}

expr_type(expr("a"))
expr_type(expr(x))
expr_type(expr(f(1, 2)))
```

```{r}
switch_expr <- function(x, ...) {
  switch(expr_type(x),
    ...,
    stop("Don't know how to handle type ", typeof(x), call. = FALSE)
  )
}
```

```{r}
logical_abbr_rec <- function(x) {
  switch_expr(x,
    # Base cases
    constant = FALSE,
    symbol = as_string(x) %in% c("F", "T"),

    # Recursive cases
    call = ,
    pairlist = purrr::some(x, logical_abbr_rec)
  )
}

logical_abbr <- function(x) {
  logical_abbr_rec(enexpr(x))
}
```

```{r}
flat_map_chr <- function(.x, .f, ...) {
  purrr::flatten_chr(purrr::map(.x, .f, ...))
}
```

Modified for problem 3.
```{r}
assignment_fns <- sapply(paste("package", c("base", "stats", "methods"),
  sep
  = ":"
), ls, pattern = "[a-zA-Z_.]+<-$") |>
  unlist(use.names = FALSE) |>
  gsub(pattern = "<-$", replacement = "")

find_assign <- function(x) unique(find_assign_rec(enexpr(x)))

find_assign_call <- function(x) {
  if (is_call(x, "<-") && is_symbol(x[[2]])) {
    lhs <- as_string(x[[2]])
    children <- as.list(x)[-1]
  } else if (is_call(x, "<-") && is.call(x[[2]]) &&
    is_symbol(x[[c(2, 1)]]) && as_string(x[[c(2, 1)]]) %in% assignment_fns &&
    is_symbol(x[[3]])) {
    lhs <- as_string(x[[3]])
    children <- as.list(x)[-c(1, 2)]
  } else {
    lhs <- character()
    children <- as.list(x)
  }

  c(lhs, flat_map_chr(children, find_assign_rec))
}

find_assign_rec <- function(x) {
  switch_expr(x,
    # Base cases
    constant = ,
    symbol = character(),

    # Recursive cases
    pairlist = flat_map_chr(x, find_assign_rec),
    call = find_assign_call(x)
  )
}

find_assign(a <- b <- c)
find_assign(names(x) <- "b")
find_assign(names(x) <- b <- "c")
```

## 1, 2. 

Seems to work...
```{r}
logical_abbr_rec <- function(x) {
  switch_expr(x,
    # Base cases
    constant = FALSE,
    symbol = as_string(x) %in% c("F", "T"),

    # Recursive cases
    closure = logical_abbr_rec(c(formals(x), body(x))),
    call = ,
    pairlist = purrr::some(if (is.symbol(x[[1]]) &&
      as_string(x[[1]]) %in% c("T", "F")) {
      x[-1]
    } else {
      x
    }, logical_abbr_rec)
  )
}
logical_abbr(T(1))
logical_abbr(T(1, 2, T))
logical_abbr(function(x = TRUE) {
  g(x + T)
})
logical_abbr(function(x = TRUE) {
  g(x + T(1))
})
logical_abbr(function(x = TRUE) {
  g(x + T(1) + T)
})
```

## 3. 

See `find_assign` modified above.

## 4. 

```{r}
find_fn <- function(x, target) {
  x <- enexpr(x)
  target <- deparse(enexpr(target))
  tryCatch(
    {
      match.fun(target)
      find_fn_rec(x, target)
    },
    error = function(e) stop(e)
  )
}

find_fn_rec <- function(x, target) {
  switch_expr(x,
    # Base cases
    constant = ,
    symbol = character(),

    # Recursive cases
    pairlist = flat_map_chr(x, find_fn_rec, target = target),
    call = {
      if (name <- as_string(x[[1]]) == target) {
        deparse(x)
      } else {
        flat_map_chr(x, find_fn_rec, target = target)
      }
    }
  )
}
find_fn(sum(1:10, sum(mtcars)), sum)
find_fn(
  {
    sum(1:10)
    x + sum
    sum()
  },
  sum
)
find_fn(lm(cyl ~ disp, data = mtcars), mean)
```

# 19. Quasiquotation

# Motivation 

## 1. 

Quoted:

* `MASS`
* `cyl == 4`
* `sum(vs)`

Unquoted:

* `mtcars`
* `mtcars2$am`
* `mtcars2` (twice)

## 2. 

Quoted: 

* `dplyr`
* `ggplot2`
* `cyl`
* `mean(mpg)`
* `cyl` and `mean` in the second call

Unquoted: 
* 
* `mtcars`
* `mpg`
* `aes(cyl, mean)`


# Quoting 

## 1. 

It's just an `enexpr` wrapper.

## 2.

Only `f2` substitutes its arguments into the expression. `f1` just quotes the expression.

## 3. 

The "arg must be a symbol" error.
```{r, error = TRUE}
enexpr(x + y)
enexpr(missing_arg())
```

## 4.

`exprs(a)` is an unnamed list with the symbol `a`, `exprs(a=)` a named list of the missing symbol.

## 5.

`exprs` offers options to force naming and ignore empty arguments.

## 6.

```{r}
substitute(x + y)

substitute(x + y, env = list(x = 4))

foo <- function(x) {
  substitute(x)
}
foo(7)
```

# Unquoting

## 1.

I had to look up the correct way to parenthesize an unquoted expression component.

```{r}
library(rlang)
xy <- expr(x + y)
xz <- expr(x + z)
yz <- expr(y + z)
abc <- exprs(a, b, c)

expr(!!xy / !!xz)
expr(-(!!xz)^!!yz)
expr(((!!xy)) + !!yz - !!xy)
expr(atan2(!!xy, !!yz))
expr(sum(!!xy, !!xy, !!yz))
expr(sum(!!!abc))
expr(mean(c(!!!abc), na.rm = TRUE))
expr(foo(!!!exprs(a = !!xy, b = !!yz)))
```

## 2.

`a`, which doesn't force the `:` call ahead of time,
is more natural, since `1:10` is only evaluated after `mean` is called.
```{r}
(a <- expr(mean(1:10)))
(b <- expr(mean(!!(1:10))))
identical(a, b)

expr_print(a)
expr_print(b)
```
# `...` 

## 1.

This implementation takes function arguments via dots and captures them using `list2`, which enables values to be defined with expressions in the caller environment, e.g. `!!key := val`. It allows the user to pass arguments to the function via a mix of standard-evaluated arguments, unquoted expressions, and unquoted lists of expressions, and control the evaluation environment.

## 2. 

`interaction` uses the simple `list(...)` and iterates over the result. `par` does the same, but substitutes the global variables `.Pars` if none are passed. Otherwise, it simplifies args to a character vector if possible, then simplifies if there is only one argument.
`expand.grid` captures the dots and returns an empty data frame if none are present. It then processes the args list in sequence, generating the expanded vectors.

## 3.

`x=10` gets matched to the `x` argument, shoving the vector intended to match `x` into `...`.
```{r}
set_attr <- function(x, ...) {
  attrs <- rlang::list2(...)
  attributes(x) <- attrs
  x
}
tryCatch(set_attr(1:10, x = 10), error = function(e) message(as.character(e)))
```

# Case Studies 

## 1. 

The first approach splices the two arguments into an `x + y` template, while the second uses `call2` to generate the reductions. I think the first form makes its intent clearer but is less "clean."

```{r}
library(purrr)
linear <- function(var, val, use_call2 = FALSE) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)
  if (use_call2) {
    reduce(summands, call2, .fn = "+")
  } else {
    reduce(summands, ~ expr(!!.x + !!.y))
  }
}
linear(x, c(10, 5, -4))
linear(x, c(10, 5, -4), use_call2 = TRUE)
```

## 2.

```{r}
bc <- function(lambda) {
  if (lambda == 0) {
    function(x) log(x)
  } else {
    function(x) (x^lambda - 1) / lambda
  }
}

bc2 <- function(lambda) {
  if (lambda == 0) {
    new_function(exprs(x = ), expr(log(x)))
  } else {
    new_function(exprs(x = ), expr((x^!!exec(`-`, lambda, 1)) / !!lambda))
  }
}
```

## 3.

```{r}
compose <- function(f, g) {
  function(...) f(g(...))
}

compose2 <- function(f, g) {
  f <- ensym(f)
  g <- ensym(g)
  new_function(exprs(... = ), expr((!!f)((!!g)(...))))
}
compose2(f, g)
```

# 20. Evaluation

## 1. 

`source` defaults to the global environment. `local = TRUE` uses the global environment, and an environment name can also be supplied.

## 2.

Outer `eval` returns a value, outer `expr` an expression.
So 4, 4, and a hideous quoted expression.

## 3.

```{r}
x <- new.env()
x$y <- 5
y <- 7

get2 <- function(name, env) {
  name <- sym(name)
  eval(name, envir = env)
}

assign2 <- function(name, value, env) {
  name <- sym(name)
  exp <- expr(`<-`(!!name, !!value))
  eval(exp, env)
}
get2("y", x)
get2("y", globalenv())
assign2("z", 7, x)
x$z
```

```{r}
source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- lapply(exprs, eval, env)
  invisible(lapply(res, print))
}
source2(here::here("misc/source_file.R"))

test <- function() {
  test2 <<- function() x
}
```

## 5.

`expr` is evaluated in `envir`, by default an empty child of the current, and wrapped in a `substitute` call. This call is then evaluated in the caller environment. 


# Quosures 

## 1.

1, 11, 111.

```{r}
q1 <- new_quosure(expr(x), env(x = 1))
q1
#> <quosure>
#> expr: ^x
#> env:  0x7fac62d19130

q2 <- new_quosure(expr(x + !!q1), env(x = 10))
q2
#> <quosure>
#> expr: ^x + (^x)
#> env:  0x7fac62e35a98

q3 <- new_quosure(expr(x + !!q2), env(x = 100))
q3
#> <quosure>
#> expr: ^x + (^x + (^x))
#> env:  0x7fac6302feb0

eval_tidy(q1)
eval_tidy(q2)
eval_tidy(q3)
```

## 2.

```{r}
enenv <- function(arg) {
  arg <- enquo(arg)
  quo_get_env(arg)
}
x <- 3
enenv(x)
with(
  mtcars,
  enenv(cyl)
)
```

# Data Masks 

## 1.

As in the example, it's possible for 
a variable to be defined based on an earlier modification of the same variable
passed via `...`. That makes vectorization impossible.

## 2.

This version inlines `rows` into the call to `[` and then evaluates it in the data mask instead of evaluating `rows` in the data mask first and calling `[` directly. 

## 3.

Unquoting `.na_last` prevents the unlikely case of  a variable by that name masking the argument.

```{r}
arrange2 <- function(.df, ..., .na.last = TRUE) {
  # quote expressions
  args <- enquos(...)
  # inline into call
  order_call <- expr(order(!!!args, na.last = !!.na.last))
  # evaluate, raising error if order is somehow the wrong length
  ord <- eval_tidy(order_call, .df)
  stopifnot(length(ord) == nrow(.df))
  # reorder
  .df[ord, , drop = FALSE]
}
```

# Using Tidy Evaluation

## 1.

The use of `$` with an unquoted symbol is harder to understand but 
avoids the need to convert the symbol to a string.

# Base Evaluation 

## 1.

In the caller environment, `data` is the name of a function, not a data frame; it only designates a data frame in the execution environment. Unquoting `data` would fix this.

## 2.

I take the factory function approach.
```{r}
auto_lm <- function(resp, .data) {
  resp <- ensym(resp)
  .data <- ensym(.data)
  formula <-
    expr(!!resp ~ !!terms)
  lm_call <- expr(lm(!!formula, data = !!.data))
  new_function(exprs(terms = ), expr({
    terms <- enexpr(terms)
    lm_call <- expr(lm(!!resp ~ !!quote(!!terms), data = !!.data))
    eval(lm_call)
  }))
}

test <- auto_lm(mpg, mtcars)
test(disp * cyl)
```

## 3.

This results in some kludgy code but evades the problem of
`data` not existing in the caller environment.
```{r}
resample <- function(df, n) {
  idx <- sample(nrow(df), n, replace = TRUE)
  df[idx, , drop = FALSE]
}

resample_lm2 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_env <- env(env, resample_data = resample_data)
  lm_call <- expr(lm(!!formula, data = data[sample(nrow(data),
    replace = TRUE
  ), , drop = FALSE]))
  expr_print(lm_call)
  eval(lm_call)
}
df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))
resample_lm2(y ~ x, data = df)
```
