---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "Chapter 13: S3"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

# Basics

## 1.

`t.test` is the generic for the `t.test` function, which dispatches to the appropriate method.
`t.data.frame` is the method of `t` for the `data.frame` class (it just coerces to matrix and invokes `NextMethod`).

## 2.

`data.frame` and many of the `as.*` and `is.*` family are major offenders.

## 3.

It coerces a data frame to a data frame, which means stripping classes inherited before `data.frame` and optionally adding row names. The overuse of dot separators makes it hard to understand that this
is an S3 method, a problem that could have been solved by using snake case instead.

## 4.


The first dispatches to `mean.Date`, which coerces a date to integer, calls `mean` on it, and coerces back to `Date`.
The second coerces to integer beforehand without doing this.

```{r}
set.seed(1014)
some_days <- as.Date("2017-01-31") + sample(10, 5)

mean(some_days)
#> [1] "2017-02-06"
mean(unclass(some_days))
#> [1] 17203
```

## 5.

It generates the ECDF function for a given vector. The object contains the function and preserves the call used to create it.
```{r}
x <- ecdf(rpois(100, 10))
x
#> Empirical CDF
#> Call: ecdf(rpois(100, 10))
#>  x[1:18] =  2,  3,  4,  ..., 2e+01, 2e+01
```

## 6.


A `table` object is an atomic vector array.
Like arrays in general, it has dimensions and a `dimnames` attribute. The class is mostly used for its print method.

```{r}
x <- table(rpois(100, 5))
x
```

# Classes

I elect not to enforce unique names, and I generate default column names the same way as `data.frame`.

## 1.

```{r}
data.frame2 <- function(..., .row_names = NULL) {
  dots <- list(...)
  l_dots <- length(dots)
  if (l_dots == 0) {
    return(structure(list(),
      class = "data.frame",
      row.names = make_row_names(.row_names, 0)
    ))
  }
  dots_names <- names(dots)
  has_names <- !is.null(dots_names)
  col_names <- vector("character", l_dots)
  col_data <- vector("list", l_dots)
  for (i in seq_along(dots)) {
    el <- dots[[i]]
    el_name <- dots_names[[i]]
    if (length(dim(el)) > 1) { # Data frame or array
      if (is.array(el)) el <- as.data.frame(el)
      el_rows <- nrow(el)
      # Ignore element name and instead combine with existing column names
      if (has_names) {
        col_names[[i]] <- ""
        if (el_name != "") {
          colnames(el) <- paste(el_name, colnames(el,
            do.NULL = TRUE,
            prefix = "V"
          ),
          sep
          = "_"
          )
        }
      }
    } else if (!is.array(el) &&
      is.vector(el)) { # Vector case
      el_rows <- length(el)
      el <- list(el) # Ensure correct concatenation
      if (has_names && el_name != "") {
        col_names[[i]] <- el_name
      }
    } else {
      stop("Cannot coerce object")
    }

    if (i == 1) {
      n_rows <- el_rows
    } else if (n_rows != el_rows) {
      stop("Number of rows mismatch")
    }
    col_data[[i]] <- el
  }

  .row_names <- make_row_names(.row_names, n_rows)

  # Supply default column names for unnamed arguments
  if (has_names) {
    unnamed <- is.na(col_names)
    col_names[unnamed] <- paste0("V", seq_along(unnamed))
    names(col_data) <- col_names
  }
  full_data <- do.call(c, col_data)
  dnn <- c(.row_names, names(full_data))
  # class(full_data) <- "data.frame"

  structure(
    .Data = full_data, class = c("data.frame"),
    row.names = .row_names
  )
  # dim = c(n_rows, length(full_data))
  # )
}

# Check row names and create if necessary
make_row_names <- function(rn, n_rows) {
  if (!is.null(rn)) {
    if (length(rn) != n_rows) {
      stop("Length of row names does not match data length")
    }
    if (anyDuplicated(rn)) stop("Duplicate row names")
  } else {
    rn <- seq_len(n_rows)
  }
  as.character(rn)
}

library(testthat)

data.frame2(mtcars, x = mtcars$cyl)
data.frame2(mpg = mtcars$mpg)
data.frame2(mpg = mtcars$mpg, cyl = mtcars$cyl)
data.frame2(unname(as.matrix(mtcars)))
data.frame2(unname(as.matrix(mtcars)), .row_names = rownames(mtcars))
data.frame2(mtcars$cyl, y = as.matrix(mtcars$disp), mtcars[4:6], .row_names = rownames(mtcars))
data.frame2()

expect_error(data.frame2(NULL))
expect_error(data.frame2(mtcars$cyl, iris$Species))
```

## 2.

I replicate the base behavior of replacing values absent from the levels with `NA` and excluding `NA` from the levels.
```{r}
new_factor <- function(x = integer(), levels = character(), contr = NULL, ...) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))
  levels <- levels[!is.na(levels)]

  nlevel <- length(unique(levels))
  out_call <-
    if (!is.null(contr)) {
      if (is.character(contr)) contr <- match.fun(contr)
      if (is.function(contr)) {
        contr <- contr(nlevel, ...)
      }
      if (!is.matrix(contr)) stop("Contrasts must be a matrix")
      if (nrow(contr) != nlevel) {
        stop("If supplied, contrasts must have one row for each level of x")
      }
    }
  out <- structure(
    x,
    levels = levels,
    class = "factor"
  )
  if (!is.null(contr)) contrasts(out) <- contr
  out
}

validate_factor <- function(x) {
  values <- unclass(x)
  levels <- attr(x, "levels")

  if (!all(values[!is.na(values)] > 0)) {
    stop(
      "All `x` values must be greater than zero",
      call. = FALSE
    )
  }


  if (length(levels) < max(values, na.rm = TRUE)) {
    stop(
      "There must be at least as many `levels` as possible values in `x`",
      call. = FALSE
    )
  }

  x
}


factor <- function(x = character(), levels = unique(x), contr = NULL, ...) {
  levels <- as.character(levels)
  ind <- match(x, levels)
  validate_factor(new_factor(ind, levels, contr = contr, ...))
}

factor(c("a", "a", "b"))
factor(1:3)
factor(1:3, levels = c(1, 3))
factor(1:3, levels = "a")
factor(1:3, contr = "contr.helmert")
expect_error(factor(1:3, levels = 1, contr = "contr.poly"))
expect_error(factor(1:3, contr = 1:5))
#> [1] a a b
#> Levels: a b
```

## 3.

The base `factor` has the additional features of mapping different labels to the same levels and ordering the factors. More
saliently, it assigns values that do not appear in the levels `NA` instead of throwing an error.

## 4.

Contrasts refer to different ways of encoding categorical variables in models in
order numerically express the effects of different levels. Above, I rewrote
`factor` to use this attribute if supplied.


## 5.

The validator should confirm that inputs are integer vectors whose elements are all in [1, 3899],  the range of valid Roman numerals, or character vectors of such valid Roman numerals. A constructor would then convert the input to integer, if necessary, then just set its class to `roman`, enabling the specialized methods to do their work.




# Generics and Methods

## 1.

It works correctly because `UseMethod` ultimately dispatches to `t.default`, since there is no `test` method for `t`.
```{r}
library(sloop)

x <- structure(1:10, class = "test")
t(x)
```

## 2.


```{r}
s3_methods_class("table")
```

## 3.


```{r}
s3_methods_class("ecdf")
```

## 4.

`print`, naturally
```{r}
generics <- lsf.str("package:base")
generics <- generics[sapply(generics, isS3stdGeneric)]
names(generics) <- generics
sapply(generics, \(x) nrow(s3_methods_generic(x))) |>
  sort(decreasing = TRUE) |>
  head()
```

## 5.


```{r}
g <- function(x) {
  x <- 10
  y <- 10
  UseMethod("g")
}
g.default <- function(x) c(x = x, y = y)

x <- 1
y <- 1
g(x)
#>  x  y
#>  1 10
```

`UseMethod` constructs a call by matching arguments in the generic's execution environment _as they came in_, forwarding them, then matching arguments defined in the execution environment. These are then forwarded to the method that is matched. So the redefinition of `x` is ignored because only the value passed is read.
This ignores the usual rule for argument lookup.

`NextMethod` skips the first method matched by `UseMethod` and continues searching, potentially to internal generics. Arguments are passed as promises to evaluate in the caller environment of NextMethod. This makes it possible to force use of a default or internal method by placing a call to `NextMethod` in a class-specific method.

## 6.

The answer differs for the many different methods implemented.

# Object Styles

## 1.

Vector : `factor`, `as.Date`, `ordered`

Record: `as.POSIXct`

Data frame: arguably `table` is a generalization

Scalar: `lm`, `ecdf`, `I`

## 2.

`lm` is a list of information related to the model. First, the call should be captured and stored. If requested, data like model weights or the QR matrix used in model fitting should be retained. Once the model is fitted, the following should be recorded for each observation:

* Observed response
* Fitted value
* Residual

Beyond that, the model matrix, and some details like factor contrasts and unique `x` levels, have to be stored as well.

# Inheritance

```{r}
new_secret <- function(x, ..., class = character()) {
  stopifnot(is.double(x))

  structure(
    x,
    ...,
    class = c(class, "secret")
  )
}
new_supersecret <- function(x) {
  new_secret(x, class = "supersecret")
}

print.supersecret <- function(x, ...) {
  print(rep("xxxxx", length(x)))
  invisible(x)
}

x2 <- new_supersecret(c(15, 1, 456))
x2

x <- structure(1:10, class = "test")
t(x)
```

## 1.

`[.Date` delegates to `NextMethod`, passing the
most specific class of the argument `x` as
determined by `oldClass`. This means a `Date` subclass `foo` is dispatched to `[.foo`.


```{r}
library(sloop)
s3_methods_generic("[")
```

## 2.

It looks like `POSIXct` methods are more verbose
and do more elaborate checking involving timezones.
`print` is naturally the same for both.

## 3.

`generic2` dispatches on the class of `x`. `generic2.b` is called first, the class reassignment is ignored, then `NextMethod` dispatches to `generic.a2`.
```{r}
generic2 <- function(x) UseMethod("generic2")
generic2.a1 <- function(x) "a1"
generic2.a2 <- function(x) "a2"
generic2.b <- function(x) {
  class(x) <- "a1"
  NextMethod()
}

generic2(structure(list(), class = c("b", "a2")))
```

# Dispatch Details


## 1.

Internal methods dispatch only on implicit class (what `1:5` has), not explicit
class set by `class`.

## 2.

`Math.data.frame` checks types before
using `lapply` to compute the operation.
`Math.difftime` records units before forwarding
to `NextMethod`. The `factor` and `PosixLT` methods
warn the user that calling them is nonsensical.
```{r}
sloop::s3_methods_generic("Math")
```

## 3.

It tracks units and throws an error for unsupported operations.
