---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "Chapter 13: S3"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

# Basics 


## 1.

`t.test` is the generic for the `t.test` function, which dispatches to the appropriate method. 
`t.data.frame` is the method of `t` for the `data.frame` class (it just coerces to matrix and invokes `NextMethod`).

## 2.

`data.frame` and many of the `as.*` and `is.*` family are major offenders.

## 3.

It coerces a data frame to a data frame, which means stripping classes inherited before `data.frame` and optionally adding row names. The overuse of dot separators makes it hard to understand that this 
is an S3 method, a problem that could have been solved by using snake case instead.

## 4.


The first dispatches to `mean.Date`, which coerces a date to integer, calls `mean` on it, and coerces back to `Date`. 
The second coerces to integer beforehand without doing this.

```{r}
set.seed(1014)
some_days <- as.Date("2017-01-31") + sample(10, 5)

mean(some_days)
#> [1] "2017-02-06"
mean(unclass(some_days))
#> [1] 17203
```

## 5.

It generates the ECDF function for a given vector. The object contains the function and preserves the call used to create it.
```{r}
x <- ecdf(rpois(100, 10))
x
#> Empirical CDF
#> Call: ecdf(rpois(100, 10))
#>  x[1:18] =  2,  3,  4,  ..., 2e+01, 2e+01
```

## 6.


A `table` object is an atomic vector array. 
Like arrays in general, it has dimensions and a `dimnames` attribute. The class is mostly used for its print method.

```{r}
x <- table(rpois(100, 5))
x
```

# Classes 

## 1.

```{r}
data.frame2 <- function(..., row_names = NULL) {
  dots <- list(...)
  l_dots <- length(dots)
  if (l_dots == 0) {
    return(structure(list(), class = "data.frame"))
  }
  dots_names <- names(dots)
  has_names <- is.null(dots_names)
  col_names <- vector("character", l_dots)
  col_data <- vector("list", l_dots)
  unnamed <- rep(FALSE, l_dots)
  for (i in seq_along(dots)) {
    el <- dots[[i]]
    if ("dim" %in% attributes(el) && dim(el) > 1) {
      if (is.array(el)) el <- as.data.frame(el)
      el_rows <- nrow(el)
      if (has_names) {
        dots_names[[i]] <- el_name
        if (el_name == "") {
          unnamed[[i]] <- TRUE
        } else {
          colnames(el) <- paste(el_name, colnames(el), sep = "_")
        }
      }
    } else {
      el_rows <- length(el)
      if (has_names && (el_name <- dots_names[[i]]) == "") {
        unnamed[[i]] <- TRUE
      } else {
        col_names[[i]] <- el_name
      }
    }
    if (i == 1) {
      n_rows <- el_rows
    } else if (n_rows != el_rows) {
      stop("Number of rows mismatch")
    }
    col_data[[i]] <- el
  }
  # Validate row names (same length as data, unique)
  if (!is.null(row_names)) {
    if (!is.character(row_names)) row_names <- as.character(row_names)
    if (length(row_names) != n_rows) stop("Length of row names does not match data length")
    if (anyDuplicated(row_names)) stop("Duplicate row names")
  }
  # Supply default column names for unnamed arguments
  col_names[unnamed] <- paste0("V", seq_along(unnamed))
  names(col_data) <- el_names
  full_data <- do.call(c, col_data)

  structure(full_data, class = c("list", "data.frame"), row.names = row_names)
}
```

## 2. 

I replicate the base behavior of replacing values absent from the levels with `NA` and excluding `NA` from the levels.
```{r}
new_factor <- function(x = integer(), levels = character()) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))
  levels <- levels[!is.na(levels)]

  structure(
    x,
    levels = levels,
    class = "factor"
  )
}

validate_factor <- function(x) {
  values <- unclass(x)
  levels <- attr(x, "levels")

  if (!all(values[!is.na(values)] > 0)) {
    stop(
      "All `x` values must be greater than zero",
      call. = FALSE
    )
  }


  if (length(levels) < max(values, na.rm = TRUE)) {
    stop(
      "There must be at least as many `levels` as possible values in `x`",
      call. = FALSE
    )
  }

  x
}


factor <- function(x = character(), levels = unique(x)) {
  levels <- as.character(levels)
  ind <- match(x, levels)
  validate_factor(new_factor(ind, levels))
}

factor(c("a", "a", "b"))
factor(1:3)
factor(1:3, levels = c(1, 3))
factor(1:3, levels = "a")
#> [1] a a b
#> Levels: a b
```

## 3. 

The base `factor` has the additional features of mapping different labels to the same levels and ordering the factors. More 
saliently, it assigns values that do not appear in the levels `NA` instead of throwing an error.

## 4. 


## 5. 


## 6. 

The validator should confirm that inputs are integer vectors whose elements are all in [1, 3899],  the range of valid Roman numerals, or character vectors of such valid Roman numerals. A constructor would then convert the input to integer, if necessary, then just set its class to `roman`, enabling the specialized methods to do their work. 


`UseMethod` constructs a call by matching arguments in the generic's execution environment _as they came in_, forwarding them, then matching arguments defined in the execution environment. These are then forwarded to the method that is matched. So the redefinition of `x` is ignored because only the value passed is read.


# Generics and Methods 

#TODO 

# Object Styles 

#TODO 

# Inheritance
```{r}
g <- function(x) {
  x <- 10
  y <- 10
  UseMethod("g")
}
g.default <- function(x) c(x = x, y = y)

x <- 1
y <- 1
g(x)
#>  x  y
#>  1 10
```

`NextMethod` skips the first method matched by `UseMethod` and continues searching, potentially to internal generics. Arguments are passed as promises to evaluate in the caller environment of NextMethod. This makes it possible to force use of a default or internal method by placing a call to `NextMethod` in a class-specific method.
```{r}
new_secret <- function(x, ..., class = character()) {
  stopifnot(is.double(x))

  structure(
    x,
    ...,
    class = c(class, "secret")
  )
}
new_supersecret <- function(x) {
  new_secret(x, class = "supersecret")
}

print.supersecret <- function(x, ...) {
  print(rep("xxxxx", length(x)))
  invisible(x)
}

x2 <- new_supersecret(c(15, 1, 456))
x2

x <- structure(1:10, class = "test")
t(x)
```

## 1.

`[.Date` delegates to `NextMethod`, passing the 
most specific class of the argument `x` as 
determined by `oldClass`. This means a `Date` subclass `foo` is dispatched to `[.foo`.


```{r}
library(sloop)
s3_methods_generic("[")
```

## 2. 

It looks like `POSIXct` methods are more verbose 
and do more elaborate checking involving timezones.
`print` is naturally the same for both.

## 3.

`generic2` dispatches on the class of `x`. `generic2.b` is called first, the class reassignment is ignored, then `NextMethod` dispatches to `generic.a2`.
```{r}
generic2 <- function(x) UseMethod("generic2")
generic2.a1 <- function(x) "a1"
generic2.a2 <- function(x) "a2"
generic2.b <- function(x) {
  class(x) <- "a1"
  NextMethod()
}

generic2(structure(list(), class = c("b", "a2")))
```

# Dispatch Details 


## 1. 

Internal methods dispatch only on implicit class (what `1:5` has), not explicit 
class set by `class`.

## 2.

`Math.data.frame` checks types before 
using `lapply` to compute the operation. 
`Math.difftime` records units before forwarding 
to `NextMethod`. The `factor` and `PosixLT` methods 
warn the user that calling them is nonsensical.
```{r}
sloop::s3_methods_generic("Math")
```

## 3. 

It tracks units and throws an error for unsupported operations.
